---
title: "Lab1"
author: "William Kett"
date: "1/13/2020"
output: 
  html_document:
    toc: yes #Table of contents
    toc_float: yes #float allows the table of contents to "move around" so it can be seen more easily
    theme: darkly
    highlights: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Lab 1

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
ddt = read.csv("DDT.csv") #ddt object now holds the output from the read.csv function applied to the .csv file
head(ddt) #gives the first six lines (by defualt) of the ddt file

```


## Task 3

1) What is the qualitative data?
The qualitative variables are: River and Species

2) What is the Quantitative data?
Mile, Length, Weight, and DDT (dichlordiphenyltrichloroethane or pesticides) in ppm.
A fascinating side note: Anything higher than 5 ppm means that the fish are considered contaminated.

3) How many species are in the DDT data set?
There are 3 species of fish in the DDT data set (if you look at the entire data set and not just the head).

4) Get a subset of the data with the specifications: [Species: 'LMB', and Weight > 800 mgs]


```{r}
with(ddt, ddt[SPECIES == "LMBASS" & WEIGHT > 800,],)


#Clicker question stuff. Ignore comments below!
#LENGTH = with(ddt, LENGTH)
#LENGTH
#mean(LENGTH)
#WEIGHT = with(ddt, WEIGHT)
#WEIGHT
#sd(WEIGHT)
#plot(WEIGHT, LENGTH, col = "Red")
#v = 1:20
#v = v/20
#v
#typeof(v)
```

5) Get a subset of the data with the specifications: [River: 'SCM', and DDT > 4]

```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0,],)
```


### Clicker Questions
  >
  Q1: 42.8125
  Q2: 376.5461
  Q3: 0 (No)
  Q4: v/20[last element] = 1.00

## Task 4

1) Make a table of the rivers

```{r}
rivs=with(ddt,table(RIVER))
rivs
```

2) Make a color coded bar plot of the rivers

```{r}
barplot(rivs, col = rainbow(4))
```

3) Make a table of the rivers crossed with the species of the fish

```{r}
rivsp=with(ddt,table(SPECIES, RIVER))
rivsp
```


4) Make a barplot of the rivers crossed with the species of the fish

```{r}
barplot(rivsp, col = rainbow(4))
```

## Task 5

1) Make a pie chart of the fish SPECIES

```{r}
sp = with(ddt, table(SPECIES))
pie(sp, col = 1:4)
```

2) Make a pie chart of the rivers

```{r}
rivs = with(ddt, table(RIVER))
pie(rivs)
```

## Task 6

1) Make a boxplot of the variable DDT

```{r}
ddt_data = with(ddt, DDT)
boxplot(x = ddt_data, col = "Red")
```

2) Make a boxplot of the variable WEIGHT

```{r}
wgt = with(ddt, WEIGHT)
boxplot(wgt, col = "Red", data = wgt)
boxplot(x = wgt,col = "Red")
```

3) Make a boxplot of the variable LENGTH

```{r}
len = with(ddt, LENGTH)
boxplot(len, col = "Red", data = len)
```

## Task 7

1) Make a coplot of LENGTH vs WEIGHT given RIVER

```{r}
coplot(LENGTH~WEIGHT|RIVER, data = ddt, col = "Red") #Using naming scheme Y vs X, the Length is the y-axis and Weight is the x-axis
```

2) Make a coplot of DDT vs WEIGHT given SPECIES

```{r}
coplot(DDT~WEIGHT|SPECIES, data = ddt, col = "Red")
```

## Task 8

```{r}
#1) Use ggplot to make boxplots
library(dplyr)
library(ggplot2)
# notice how the infix %>% operator works ( the pipe )

ddt %>% filter() %>% mutate(AREA = LENGTH*WEIGHT) %>% ggplot(aes(x=SPECIES,y=WEIGHT,fill=RIVER)) + geom_boxplot() + ggtitle("William Kett")

```

2) Use ggplot to make violin plots

```{r}
ddt %>% filter() %>% mutate(AREA = LENGTH*WEIGHT) %>% ggplot(aes(x=RIVER,y=LENGTH,fill=SPECIES)) + geom_violin() + ggtitle("William Kett")
```
3) use ggplot to make scatter plots

```{r}
pointcol = with(ddt, ifelse(SPECIES == "CCATFISH", "Red",         #if Species is "CCATFISH", point is Red
                            ifelse(SPECIES == "LMBASS", "Green",  #else if Species is "LMBASS", point is Green
                                   "Blue")))                      #else point it Blue
              
ddt %>% filter() %>% mutate(AREA = LENGTH*WEIGHT) %>% ggplot(aes(x=WEIGHT,y=LENGTH,fill=SPECIES)) + geom_point(col = pointcol) + ggtitle("William Kett")
```



