---
title: 'Writing Functions: Class exercise'
author: "Wayne Stewart"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment='ws#')
```

# Introduction
>
We will be making R functions in this course and in fact the first project will require you to make a number of them.

```{r,eval=FALSE}
functionname <- function(x,y,z, ...){
  R expressions
  last line released to the command line
}
```


# Task

>
You will need to make a function that takes a vector of quantitative data as input and then calculates the standard deviation.

## Specifically

Use the following formula:

$$s = \sqrt{\frac{\sum_{i=1}^n (x_i-\bar{x})^2}{n-1}}$$




1. Fill in the code below:

```{r sd}

mysd = function(x){ # x is a vector
  
}

```

2. Run the function with the following input. Notice the chunk option `eval=FALSE` -- you will need to remove this once your function is made.

```{r test1,eval=FALSE}
L=1:20
mysd(L)

```

3. Now compare with the built in `R` function `sd()` by calculating the standard deviation of `L` : that is calculate `sd(L)`. Put in R chunk below

```{r compare}
cat("Use this chunk")
```


4. This is important: Alter your function so that the last line before the `}` is a `list` function containing both the standard deviation and the variance.


```{r,}
# this is how lists work
aobject = 1:10
bobject = "Jack and the bean stalk"
list(numbers = aobject, story = bobject)
```

Invoke the adjusted function by first removing the chunk option `eval=FALSE`

```{r, eval=FALSE}
mysd(x=L)
```


5. Upload the html document to CANVAS

